---
title: "Alternate Project"
author: "Kevin Sun"
date: "`r Sys.Date()`"
output:
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message = F,warning = F)
```

# Introduction


The Titanic disaster is one of the most well-known maritime tragedies in history. In 1912, the Titanic, a British passenger liner, sank in the North Atlantic Ocean after colliding with an iceberg. The disaster resulted in the loss of more than 1,500 lives. The Titanic disaster continues to fascinate people, and there have been numerous books, films, and documentaries produced about it.

This dataset contains cleaned data on the passengers who were aboard the Titanic when it sank. The data includes information on the passengers' age, gender, ticket class, cabin class, and whether they survived the disaster. The purpose of this visualization report is to explore and analyze this dataset to gain insights into the factors that influenced the survival rate of the passengers. By analyzing the data, we hope to gain a better understanding of the Titanic disaster and its impact on the passengers who were aboard the ship.

# Data summary



The Titanic cleaned data set contains information on 891 of the 2224 passengers and crew members who were on board the Titanic. The data has been cleaned and preprocessed for analysis. It includes information on various attributes such as passenger age, gender, ticket class, cabin class, and whether the passenger survived the disaster or not.


```{r}
library(tidyverse)
library(psych)
library(randomForest)

theme_set(theme_minimal())

df = read_csv("data/train_clean.csv")

df %>% mutate(Survived = ifelse(Survived ==1, 'Survived', 
                                       'NonSurvived')) ->df

```

- PassengerId: Unique identifier for each passenger
- Survived: Whether the passenger survived or not (0 = No, 1 = Yes)
- Pclass: Ticket class (1 = 1st, 2 = 2nd, 3 = 3rd)
- Name: Passenger's name
- Sex: Passenger's sex
- Age: Passenger's age in years
- SibSp: Number of siblings/spouses aboard the Titanic
- Parch: Number of parents/children aboard the Titanic
- Ticket: Ticket number
- Fare: Passenger fare
- Cabin: Cabin number
- Embarked: Port of embarkation (C = Cherbourg, Q = Queenstown, S = Southampton)


The characteristics of the dataset affect the analysis methodology. 

- the variables which are neatural for surviving will be dropped, that is, `Cabin`, `PassengerId`, `Ticket` and `Name`.

- Luckily, there is no missing value after dropping these columns.

- Moreover, there are both categorical and numerical variables, we will take analysis on them seperately.



To summarize the dataset, we can begin by examining the basic statistics of the numerical variables (such as age, fare, etc.) using measures such as mean, median, standard deviation, and quartiles. 


```{r}
library(pander)
num_df = df %>% select(
  Age, Fare, Parch, SibSp,  Family_Size
)

cat_df = df %>%
  select(-c(Age, Fare, Parch, SibSp,  Family_Size)) %>%
  select(-c(Cabin, Name, PassengerId,Ticket))


for(col in colnames(cat_df)){
  cat_df[,col] = factor(cat_df %>% pull(col))
}


uni_cat = function(col){
  
  cat_df %>%
    group_by(!!sym(col)) %>%
    count() %>%
    ungroup() %>%
    mutate(proportion = n /sum(n) )-> ct
  
  ct %>%
    ggplot(aes(!!sym(col), proportion, fill = !!sym(col))) +
    geom_col() +
    geom_text(aes(label = n),vjust = -0.5) +
    theme(legend.position = "none") + 
    labs(title = "col distribution" %>%
           gsub('col', col, .)) ->fig
  
  return(fig)
}






num_df %>% pivot_longer(cols = num_df %>%
                          colnames()) %>%
  group_by(name) %>%
  summarise(n_levels = length(unique(value))) -> values_ct


values_ct %>% filter(n_levels <10) %>%
  pull(name) -> may_cat


num_df %>%
  describe() %>%
  select( mean:median, min:kurtosis) %>%
  pander()
  

```

Following histogram plots provide the  visualization for these fact, the distribution of age seems being symmetric, however, all other variables' distributions are quite right skewed.


```{r}
num_df %>% pivot_longer(cols = num_df %>%
                          colnames()) %>%
  ggplot(aes(y = value, fill = name)) +
  geom_histogram() +
  facet_grid(vars(name), scales = 'free') +
  theme(legend.position = "none") +
  labs(title = "Distribution of numerical features")
```

We can then create frequency tables to examine the distribution of categorical variables , however, this could be better demonstrated using bar chart, the y axis represent the proportion, while the number above the bar is the absolute number in each level, the data is not even distributed, regardless we use which categorical variable to classify the data.


```{r}
uni_cat('Pclass')
uni_cat('Embarked')
uni_cat('Sex')
uni_cat('Survived')
uni_cat('Title')
```




# Data analytics


## Bivariate analysis


Bivariate analysis is a statistical analysis method used to explore the relationship between two variables. In this case, we will use bivariate analysis to explore the relationship between various features and the outcome variable, "Survived".

For numerical features, we compare the distributions by varying the survived status, the result are expressed by side by side boxplot:

```{r}

bi_num = function(col){
  
  stat = df %>%
    group_by(Survived) %>%
    summarise(describe(!!sym(col)))  %>%
    select(Survived, mean, median, sd, skew, kurtosis)
  
  df %>%
    ggplot(aes(x = Survived, y = !!sym(col) , fill = Survived)) +
    geom_boxplot()+
    theme(legend.position = "none") +
    labs(title = "comparing col by survived" %>%
           gsub("col", col, .)) ->f
  
  print(f)
  
  return(list(s = stat, f = f))
  
  
}

tmp = bi_num('Age')
tmp$s %>% pander()
```



Based on the statistics and visualization, people in NonSurvived group is elder.


```{r}
tmp = bi_num('Fare')
tmp$s %>% pander()
```


Based on the statistics and visualization, fare in NonSurvived group is higher.


```{r}
tmp = bi_num('Parch')
tmp$s %>% pander()
```


Based on the statistics and visualization, It is unclear the relationship between Survived and Parch.

```{r}
tmp = bi_num('SibSp')
tmp$s %>% pander()
```

Based on the statistics and visualization, It is unclear the relationship between Survived and SibSp.

```{r}

tmp = bi_num('Family_Size')
tmp$s %>% pander()
```

Based on the statistics and visualization, family size in NonSurvived group is larger.


For categorical features, we compare the distributions by varying the survived status, the result are expressed by side by side bar chart:


```{r}
bi_cat = function(col){
  df %>% group_by(
    Survived, !!sym(col)
  ) %>%
    summarise(n = n()) %>%
    mutate(proportion = (n /sum(n)) %>%
             round(3)) ->ct
  
  ct %>%
    ggplot(aes(Survived, proportion, fill = !!sym(col) %>%
                 as.factor())) +
    geom_col(position = "dodge") +
    labs(title = "comparing col by survived" %>%
           gsub("col", col, .), fill = col) ->f
  
  return(list(s = ct, f = f))
  
}


bi_cat('Pclass')$f
```



The proportion of class 3 in NonSurvived group is quite higher.




```{r}
bi_cat('Embarked')$f
```

The proportion of Embarked port C in Survived group is quite higher. The proportion of Embarked port S in NonSurvived group is quite higher.

```{r}
bi_cat('Sex')$f
```


The proportion of males in NonSurvived group is quite higher.


```{r}
bi_cat('Title')$f
```


The proportion of people with Title `Mrs` in Survived group is quite higher, as well as the propotion of people with title `Master`.


We also examine the association between some numerical encoded variables and the outcome variable in the way we convert them into categorical variables. More or less, this convertion may be reasonable, for example, it is unclear the relationship between Survived and SibSp, however, in case we treat SibSp as categorical, the proportion single SibSp in Survived group is quite higher.

```{r}
bi_cat('Parch')$f
```


```{r}
bi_cat('SibSp')$f

```


```{r}
bi_cat('Family_Size')$f
```



## Multviarte analysis

We use logistic regression and random forest classifier to determine the most impact factor on Survived.


### logistic regression

In logistic regression, we use all possible predictors in a full model, then we remove the predictors step wise based on AIC. Following table gives the estimation:

```{r}
mdf = cbind(num_df, cat_df)

mdf$Survived = as.factor(mdf$Survived)

glm(Survived ~ ., data = mdf, family = binomial()) -> lgr

step(lgr, trace = F) -> back

back %>% summary() %>%
  coef()-> backs

backs %>%
  pander()


```

Using the absolute z value as measure for impact level, following figure visualize the impact level and direction:

```{r}
backs %>%
  as.data.frame() %>%
  mutate(Features = rownames(backs),
         abs = abs(`z value`))  %>%
  filter(`Pr(>|z|)` < 0.05) %>%
  ggplot(aes(`z value` ,Features %>%
               reorder(abs), fill = `z value`)) +
  geom_col() +theme(legend.position = "none" )+
  labs(title = "Impact of Features (Logistic Regression)",  y = 'Features')
```


### Random Forest

Also, we use random forest model to determine the rank of importance of predictors:

```{r}
randomForest(Survived ~ ., data = mdf) -> rf
importance(rf) %>% as.data.frame() %>%
  arrange(-MeanDecreaseGini) ->rf_imp
rf_imp$Variable = rownames(rf_imp)
rf_imp %>% ggplot(aes(Variable %>%reorder(-MeanDecreaseGini), 
                      MeanDecreaseGini, fill = -MeanDecreaseGini)) +
  geom_col() +
  theme(legend.position = "none" )+
  labs(title = "Importance of Features (Random Forest)",  x = 'Features')
```



# Conclusion


The backward elimination approach using logistic regression and random forest models helps us identify the most important features that predict survival on the Titanic.

According to the logistic regression model, the features that most significantly reduce the probability of survival are being in the 3rd class cabin, having a large number of siblings/spouses aboard (SibSp), and being in the 2nd class cabin. In addition, increasing age is associated with a lower probability of survival. On the other hand, having the title of Mrs., followed by the titles of Master and Miss, is associated with higher probabilities of survival.

According to the random forest model, the top three most important features in predicting survival are Title, Fare, and Gender. This implies that the passenger's title, their fare, and their gender are highly indicative of whether they are likely to survive or not.

Overall, the results of the logistic regression and random forest models suggest that certain passenger attributes, such as cabin class, age, title, fare, and gender, have a significant impact on the likelihood of survival. This information can be useful in developing a predictive model to estimate the likelihood of survival of passengers on the Titanic. It can also help us better understand the factors that influenced the survival outcomes of passengers during the disaster.




# DevOps

Using GitHub for a solo project provides several benefits, including the ability to track changes and maintain a backup of your code, build a portfolio, and collaborate with others if desired. However, there is a learning curve involved in using GitHub effectively, and it requires ongoing maintenance to keep your code up-to-date on the platform. Additionally, if you have sensitive information in your project, such as passwords or personal data, it may not be safe to store it on a public platform like GitHub.

For group projects, GitHub provides several advantages, such as enabling efficient and effective collaboration, allowing multiple users to work on the same project simultaneously, tracking changes and resolving conflicts, and helping to keep the project organized and on track. GitHub also allows you to control who has access to the project, which can be important if you need to keep sensitive information secure. However, as with solo projects, there is a learning curve involved, and merge conflicts can occur when multiple people are working on the same code. Additionally, it's important to establish guidelines around code ownership and use, which can be more complex in a group setting. Finally, GitHub can experience technical issues from time to time, which can impact productivity. Overall, the benefits of using GitHub for both solo and group projects generally outweigh the drawbacks, but it's important to carefully consider the potential challenges before deciding to use the platform.


